---
title: "Code Reste"
author: "Johanna Mehler"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Country-Specific Patterns

```{r}
# add country column to the dataframe again - but needs to have same amount of rows...
data_numeric_country <- data_numeric %>% mutate(Input.country = batches$Input.country)

country_specific_patterns <- data_numeric_country %>%
  group_by(Input.country) %>%
  summarise(across(starts_with("Answer."), sum, na.rm = TRUE))

# Preview the results for a few countries
head(country_specific_patterns)
```
Numbers BEFORE filtering, need to be adjusted

* Great Britain (gbr): High frequency of attack content (116), incitement content (66), and a notable focus on hateful content (48).
* Germany (ger): Similar to Great Britain in terms of attack content (111) but with a higher frequency of dehumanization content (49) and incitement content (72).
* India (English responses, indeng): Shows a lower frequency of attack content (59) compared to GB and Germany, with a significant focus on religion as a target (55).
* Nigeria (nig): Notable for high dehumanization content (33) and a focus on race (51) and religion (45) as targets.
* Philippines (English responses, phleng): Similar to Nigeria in terms of dehumanization content (32) and a focus on hateful content (34).


# Clustering

Since your variables are boolean, K-modes is more suitable than K-means. You can use the klaR package for K-modes:
```{r}
library(klaR) # ACHTUNG stürzt immer ab

kmodes_result <- kmodes(batches, modes = 5)

kmodes_result
```

# Filtering for language and checking country selection

```{r}
# filter: for now only English answers (because the others are not translated yet)
data_en <- data_all %>% filter(Q_Language == "EN")

# for now we are left with three countries: usa (2945), phleng (837), nig (464)
data_en %>% select(country) %>% unique()
```

For now we are left with three countries: usa (2945), phleng (837), nig (464), all together 4,246 English observations


# Missingness

Hate Definition (open_hatedefinition): 19.4 (24.1???) per cent missing. Seen that there is high missingness of nearly a quarter in my outcome variable, I want to find out more about people who did not specify their own definition of hate speech.

```{r}
# Filter rows where open_hatedefinition is missing
missing_hatedef <- data %>% 
  filter(is.na(open_hatedefinition)) 

not_missing_hatedef <- data %>% 
  filter(!is.na(open_hatedefinition)) 

# Comparing sociodemographics of those who did not answer with the rest

# Plotting for each variable
variables <- data %>% select(-open_hatedefinition) %>% colnames()

for (var in variables) {
  # Plot for missing_hatedef dataset
  p1 <- ggplot(data = missing_hatedef, aes(x = .data[[var]])) +
    geom_bar() +
    theme_minimal() +
    labs(title = paste("Missing:", var))
  
  # Plot for not_missing_hatedef dataset
  p2 <- ggplot(data = not_missing_hatedef, aes(x = .data[[var]])) +
    geom_bar() +
    theme_minimal() +
    labs(title = paste("Not Missing:", var))
  
  # Combine plots
  plot_list <- list(p1, p2)
  multiplot <- cowplot::plot_grid(plotlist = plot_list, nrow = 1)
  
  # Show the plots
  print(multiplot)
}
```

Gender: seems not to be important for this.
Age: in the age of 18-29, much less people are missing a definition -> make a more sophisticated plot out of this (with percentages only)
Minority: have provided a bit more information -> also show in plot
Academic: slightly different in non-academics -> also show in plot!


# Examining Batches

```{r}
# read in batches
Batch_1 <- read_csv("~/OneDrive/1_Hertie Studies/Thesis/2023-hate-speech-opentext/definition-classifications/Batch_394877_batch_results.csv")
Batch_2 <- read_csv("~/OneDrive/1_Hertie Studies/Thesis/2023-hate-speech-opentext/definition-classifications/Batch_395273_batch_results.csv")

# find out which annotators are represented in each batch
Batch_1 %>% select(WorkerId) %>% unique() # 2 annotators
Batch_2 %>% select(WorkerId) %>% unique() # same 2 annotators

Batch_1 %>% group_by(WorkerId) %>% count()
Batch_2 %>% group_by(WorkerId) %>% count()
```







