---
title: "Missingness Descriptives"
author: "Johanna Mehler"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true               # Enable Table of Contents
    toc_float:              # Make the TOC floating
      collapsed: false      # Do not collapse TOC by default
      smooth_scroll: true   # Smooth scrolling to sections
    number_sections: true   # Numbered sections
    theme: readable         # Use a predefined theme for styling (optional)
    highlight: tango        # Syntax highlighting style (optional)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Preparation

```{r}
library(tidyverse)
library(ggplot2)
library(here)

# load reprocessed data
data_all <- read_csv(here("data/data_clean.csv"))
data_controls <- read_csv(here("data/controls.csv"))

# join data by ResponseId
data <- inner_join(data_all, data_controls, by = "ResponseId")

# select hate definition, edu variables, and controls 
data <- data %>% select(open_hatedefinition, minority_cat, academic_status, educ_cat, gender.x, age_cat.x, polinterest_cat_3, empathy_pc_cat, exp_hate_speech_cat, exp_hostile_engagement_cat)

## More Covariates ####
## select variables of interest
#data <- data_all %>% select(open_hatedefinition, gender, age_cat, minority_cat, educ_cat, #academic_status, polinterest_cat, empathy_person, empathy_predicting, empathy_perspective, #speak_freely, speak_freely_pers, silencing_harmful, silencing_unacceptable, starts_with("exp"), #tradeoffs_speech) 
```


# Missingness

Hate Definition (open_hatedefinition): 19.4 (24.1???) per cent missing. Seen that there is high missingness of nearly a quarter in my outcome variable, I want to find out more about people who did not specify their own definition of hate speech.

## Subset data in missing and not missing the hate speech definition

```{r}
# Filter rows where open_hatedefinition is missing
missing_hatedef <- data %>% 
  filter(is.na(open_hatedefinition)) 

not_missing_hatedef <- data %>% 
  filter(!is.na(open_hatedefinition)) 

# Create vector of variable names
variables <- data %>% select(-open_hatedefinition) %>% colnames()
```

## Calculate Percentages of each value in each group

```{r}
# group "missing"

# convert to long format
missing_hatedef_long <- missing_hatedef %>% 
  pivot_longer(cols = variables, names_to = "variable", values_to = "value")

# summarize counts
missing_hatedef_summary <- missing_hatedef_long %>% group_by(variable, value) %>% count()

# calculate total of each variable
total <- missing_hatedef %>% nrow()

# add total and share of each value, and dataset variable "missing"
missing_hatedef_summary <- missing_hatedef_summary %>% mutate(
  total = total,
  share = n/total,
  dataset = "Missing"
)

# group "not missing"

# convert to long format
not_missing_hatedef_long <- not_missing_hatedef %>% 
  pivot_longer(cols = variables, names_to = "variable", values_to = "value")

# summarize counts
not_missing_hatedef_summary <- not_missing_hatedef_long %>% group_by(variable, value) %>% count()

# calculate total of each variable
total <- not_missing_hatedef %>% nrow()

# add total and share of each value, and dataset variable "not missing"
not_missing_hatedef_summary <- not_missing_hatedef_summary %>% mutate(
  total = total,
  share = n/total,
  dataset = "Not Missing"
)

# combine missing and not missing dataset summaries
missingness_sociodemographics <- rbind(missing_hatedef_summary, not_missing_hatedef_summary)
```

## Create plots

```{r, fig.width=9, fig.height=4}
# plot with geom_point
missingness_sociodemographics %>% 
  ggplot(aes(x = share, y = value)) +
  geom_point(aes(color = dataset), stat = "identity") + # Use geom_bar to create a bar plot
  geom_line(aes(group = value), color = "black", show.legend = FALSE) +
  facet_wrap(~variable, scales = "free", ncol = 3) + # Separate plots for each Region
  labs(title = "Sociodemographics of those Missing compared to existing answers", x = "Share", y = "Value")
```

Are these differences significant? 

```{r, fig.width=10, fig.height=4}
# plot without line but with confident intervals

# Add standard error and confidence interval bounds to the dataframe
missingness_sociodemographics <- missingness_sociodemographics %>%
  mutate(
    share = as.numeric(share), # Ensure that 'share' is numeric
    SE = sqrt(share * (1 - share) / total), # Standard error of the proportion
    z_value = qnorm(0.975), # z-value for a 95% CI
    lower = share - z_value * SE, # Lower bound of the CI
    upper = share + z_value * SE # Upper bound of the CI
  )

# create the plot with point ranges
ggplot(missingness_sociodemographics, aes(x = factor(value), y = share, color = dataset)) +
  geom_pointrange(aes(ymin = lower, ymax = upper), position = position_dodge(width = 0.1), fatten = 3) +
  facet_wrap(~variable, scales = "free", ncol = 3) +
  labs(title = "Sociodemographics of those Missing compared to existing answers",
       x = "Value", y = "Share") +
  scale_x_discrete(na.translate = FALSE) + # remove NA values from the x-axis
  coord_flip()
```

People who did not answer...

* have less academic education
* are more often between 18 and 29 and more rare in the age group 50-69
* ...
* ...
* ...

X, Y and Z seem to be not significantly different between the groups of missing and not missing answers.

The high missingness quota could also be traced back to the very long survey experiment and missing incentives to take a lot of time at the end to answer the questions (answering time for the open text questions was highest compared to all other items of the survey).


## Bar plot

```{r, fig.width=8, fig.height=8}
# bar plot
missingness_sociodemographics %>%
  ggplot(aes(x = value, y = share, fill = dataset)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.75)) + 
  geom_errorbar(aes(ymin = lower, ymax = upper, group = dataset),
                width = 0.3,                    # Width of the horizontal lines at the ends of error bars
                position = position_dodge(0.75), # Match the position_dodge width with geom_bar
                color = "black",                 # Color of the error bars
                size = 0.5) +                   # Size of the error bars, making them narrow
  facet_wrap(~variable, scales = "free", ncol = 2) +
  labs(title = "Missingness Sociodemographics", x = "Value", y = "Share") +
  scale_x_discrete(na.translate = FALSE) +
  coord_flip()

```
