---
title: "batches_descriptives"
author: "Johanna Mehler"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true               # Enable Table of Contents
    toc_float:              # Make the TOC floating
      collapsed: false      # Do not collapse TOC by default
      smooth_scroll: true   # Smooth scrolling to sections
    number_sections: true   # Numbered sections
    theme: readable         # Use a predefined theme for styling (optional)
    highlight: tango        # Syntax highlighting style (optional)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Libraries

```{r, include=FALSE}
library(tidyverse)
library(descr) 
library(ggplot2)
library(descr)


theme_set(theme_minimal())
```

## Read in Data

```{r}
# read in numeric data
data_numeric_filtered <- read_csv("~/OneDrive/1_Hertie Studies/Thesis/Hertie-Thesis-Mehler/data/data_numeric_filtered.csv")
data_content <- read_csv("~/OneDrive/1_Hertie Studies/Thesis/Hertie-Thesis-Mehler/data/data_content.csv")
data_other_features <- read_csv("~/OneDrive/1_Hertie Studies/Thesis/Hertie-Thesis-Mehler/data/data_other_features.csv")
data_scope <- read_csv("~/OneDrive/1_Hertie Studies/Thesis/Hertie-Thesis-Mehler/data/data_scope.csv")
data_sender <- read_csv("~/OneDrive/1_Hertie Studies/Thesis/Hertie-Thesis-Mehler/data/data_sender.csv")
data_target <- read_csv("~/OneDrive/1_Hertie Studies/Thesis/Hertie-Thesis-Mehler/data/data_target.csv")
```

# Explore Distributions

```{r}
# Count the number of 1s in each column of the dataframe
ones_count <- colSums(data_numeric_filtered == 1)

# Convert the counts to a dataframe for plotting
ones_count_df <- data.frame(variable = names(ones_count), count = ones_count) 

# Order the variables by count in descending order
ones_count_df$variable <- reorder(ones_count_df$variable, -ones_count_df$count)

# Create a bar plot with each column's number of 1s
ggplot(ones_count_df, aes(x = variable, y = count)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + # Rotate x-axis labels for better readability
  labs(title = "Counts per Classification Item", x = "Item", y = "Count")

```
```{r}
# Overall patterns: Count the frequency of TRUE values for each Answer.* column
overall_patterns <- colSums(data_numeric_filtered, na.rm = TRUE) # Assuming the first column is `Input.country`
overall_patterns_sorted <- sort(overall_patterns, decreasing = TRUE)

# Display the most common properties/content
head(overall_patterns_sorted, 15)
```

These results indicate that the most common types of content in responses involve attacks, offensive speech, and incitement, often targeting race or religion and focusing on either individuals or groups. The share of nonsense answers seems big and should be further investigated, especially in terms of a potential underlying political attitude.

## Content items

```{r}
# Count the number of 1s in each column of the dataframe
ones_count_content <- colSums(data_content == 1)

# Convert the counts to a dataframe for plotting
ones_count_content_df <- data.frame(variable = names(ones_count_content), count = ones_count_content) 

# Create a bar plot with each column's number of 1s
ggplot(ones_count_content_df, aes(x = reorder(variable, -count), y = count)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + # Rotate x-axis labels for better readability
  labs(title = "Counts per Content Item", x = "Item", y = "Count")

```

## Sender items

```{r}
# Count the number of 1s in each column of the dataframe
ones_count_sender <- colSums(data_sender == 1)

# Convert the counts to a dataframe for plotting
ones_count_sender_df <- data.frame(variable = names(ones_count_sender), count = ones_count_sender) 

# Create a bar plot with each column's number of 1s
ggplot(ones_count_sender_df, aes(x = reorder(variable, -count), y = count)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + # Rotate x-axis labels for better readability
  labs(title = "Counts per Sender Item", x = "Item", y = "Count")

```

## Target items

```{r}
# Count the number of 1s in each column of the dataframe
ones_count_target <- colSums(data_target == 1)

# Convert the counts to a dataframe for plotting
ones_count_target_df <- data.frame(variable = names(ones_count_target), count = ones_count_target) 

# Create a bar plot with each column's number of 1s
ggplot(ones_count_target_df, aes(x = reorder(variable, -count), y = count)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + # Rotate x-axis labels for better readability
  labs(title = "Counts per Target Item", x = "Item", y = "Count")

```

## Other Features of Statement

```{r}
# Count the number of 1s in each column of the dataframe
ones_count_other <- colSums(data_other_features == 1)

# Convert the counts to a dataframe for plotting
ones_count_data_other <- data.frame(variable = names(ones_count_other), count = ones_count_other) 

# Create a bar plot with each column's number of 1s
ggplot(ones_count_data_other, aes(x = reorder(variable, -count), y = count)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + # Rotate x-axis labels for better readability
  labs(title = "Number of 1s per Variable", x = "Variable", y = "Count")

```

## Scope

```{r}
# Count the number of 1s in each column of the dataframe
ones_count_scope <- colSums(data_scope == 1)

# Convert the counts to a dataframe for plotting
ones_count_scope <- data.frame(variable = names(ones_count_scope), count = ones_count_scope) 

# Create a bar plot with each column's number of 1s
ggplot(ones_count_scope, aes(x = reorder(variable, -count), y = count)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + # Rotate x-axis labels for better readability
  labs(title = "Scope", x = "Item", y = "Count")

```

# Items per person

```{r}
item_counts <- data_numeric %>% 
  # Convert TRUE/FALSE to 1/0 and sum them for each row
  mutate(TotalTrues = rowSums(across(everything()))) %>%
  # Arrange the data from most to fewest TRUEs
  arrange(desc(TotalTrues)) 

freq(item_counts$TotalTrues)
```


